### 2013-02-20 phys761/phys761assign3.tex

\documentclass{article}
\author{Ben Keller}
\title{Physics 761 Presentation: "Population III stars: hidden or disappeared?"}
\begin{document}
\maketitle
\section{Background: Metallicity \& Stellar Populations}
\section{The Paper's Methodology}
This paper tried to answer the question of how late could Population III stars
could still form using a series of numerical simulations.  The authors used 
smoothed particle hydrodynamics (SPH) to simulate a comoving volume of VOLUME 
Mpc$^3$.  This volume was filled with $2*256^3$ dark matter and baryonic 
particles ($256^3$ of each).  These particles had a mass 
\section{The Paper's Results}
\section{Conclusions}
\end{document}