### 2013-02-21 phys785/paper/paper.tex
Added some quick outline sections

\documentclass[12pt,usenatbib]{article}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{geometry}
\usepackage{gensymb}
%\usepackage{epsfig}
\doublespacing

\geometry{
  body={5in, 8.5in},
left=1.25in,
right=1.25in,
  top=1.5in
}

\begin{document}

\title{Physics 785 Term Paper: Turbulence in Molecular Clouds: Origins \&
Impacts}
\author{Ben Keller}
\maketitle
\newpage
\begin{abstract}
	Turbulence yo.
\end{abstract}
\section{Introduction}
\section{Observational Evidence}
\section{Origins of Turbulence}
\section{Impact of Turbulence on Cloud Structure \& Evolution}
\section{Conclusion}
\bibliographystyle{mn2e}
\bibliography{references}
\end{document}